---
title: "R_vdeq_toregram"
author: "Michael McManus, US EPA/ORD"
date: "12/03/2024"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline

Our exploratory spatial data analysis (ESDA) continues by examining spatial variation now over stream network distances, instead of only Euclidean distance. 

## Libraries and Refernces
```{r libraries}
library(tidyverse)
library(lubridate)
library(sf)
library(mapview)
library(leaflet)
library(leafpop) # for popups in mapview
library(gstat) # for semivariograms
library(lattice) # for random semivariogram plots
library(spmodel) # for spatial modeling
library(scales) # comma instead of scientific notation
library(plotly) # interactive plots
library(spmodel) # has empirical semivariogram (esv) function
library(SSN2) # for spatial stream network (SSN) objects
library(janitor) # clean_names function

```

## SSN Import

A spatial stream network object is needed to evaluate semivariance in the response variable as a function of Euclidean and stream network distances, flow-connected and flow-unconnected. 

Ideally for an SSN analysis, the data set has these 3 characteristics:
* Sample size:  minimum of 50, maximum of 2,000
* Spatial Configuration:  headwaters to outlet
* Spatial Clustering:  around confluences.

```{r ssn_import}
j_ssn1a <- SSN2::ssn_import("ssn_object/James_071024_pluspreds.ssn", predpts = "sites")
class(j_ssn1a)

names(j_ssn1a)
names(j_ssn1a$obs)

# for consistency I am going to pull the obs from the ssn and apply clean_names so all variable names are in lower case

DFobs <- SSN2::ssn_get_data(j_ssn1a)
DFobs <- clean_names(DFobs)
names(DFobs)

# this shows that DFobs is both an sf and data.frame
class(DFobs)

# note ssn_put_data requires sf object and SSN2 object
# this is putting cleaned names of DFobs back into SSN2 object
j_ssn1a <-  SSN2::ssn_put_data(DFobs,j_ssn1a)

```

## Create Distance Matrix
SSN2 creates a distance folder where distance R object stored. Look where the SSN object is located in File Explorer. In our case we are looking for the James_071024_pluspreds.ssn folder and within that is the distance folder. The creation of the distance matrix to calculate all three distant types only needs to be done once as long as the geography of points (observation or sites, and prediction points) and flowlines (edges) have not been altered. The function has already been run so it is commented out. Note that this distance matrix only contains distances for obs. When we go to make predictions with an SSN model we need to make sure and run the distance matrix and specify that is also to include distances for preds.
```{r distance_matrix}
## Create distance matrices for observed sites
# SSN2::ssn_create_distmat(j_ssn2, overwrite = TRUE)
```

## Torgegram
When semivariance is plotted as a function of stream network distances the plot is called a Torgegram. Why might the flow-connected Toregram look so wonky? If you examine the flow-connected data frame created below you will see the incredibly small sample size for the 15 plotted points. One typically wants the number of pairs of points (np) to be at least 30 for each semivariance point. The wonky pattern associated with the small sample size of the flow-connected Torgeram is why that distance, and its associated tail-up covariance component, was not considered in the SSN modeling.
```{r torgegram}
names(j_ssn1a)
summary(j_ssn1a)

ztg <- SSN2::Torgegram(vscivcpmi ~ 1, j_ssn1a, type = c("flowcon", "flowuncon", "euclid"))
plot(ztg, main = "VSCI")
names(ztg)

View(ztg$euclid)

torg_eu <- ztg[["euclid"]]
names(torg_eu)
class(torg_eu)
ggplot(torg_eu, aes(x=dist, y=gamma,size=np)) + geom_point() + ggtitle("VSCI Euclidean Semivariogram")

torg_fc <- ztg[["flowcon"]]
names(torg_fc)
class(torg_fc)
ggplot(torg_fc, aes(x=dist, y=gamma,size=np)) + geom_point() + ggtitle("VSCI Flow-Connected Torgegram")

torg_fu <- ztg[["flowuncon"]]
names(torg_fu)
class(torg_fu)
ggplot(torg_fu, aes(x=dist, y=gamma,size=np)) + geom_point() + ggtitle("VSCI Flow-Unconnected Torgegram")

# specifying separate = TRUE requires hitting the return bar to see each Torgegram. If trying to run all the code chunks from top to bottom at once, then comment this line out otherwise it will error.

# plot(ztg, separate = TRUE, main = "VSCI")

```

The Euclidean semivariogram suggest some spatial structure at distances up to 75 km apart. As noted, the flow-connected Torgegram does not look spatially interpretable. How did the manner in which the sites were selected likely affect the wonky flow-connected pattern? The flow-unconnected Torgegram suggests some spatial structure as well.

We have evidence of spatial autocorrelation in VSCI. But, what we want to know is after we have accounted for variation in VSCI by modeling using covariates, is there leftover, or residual, spatial autocorrelation. If that's the case, then we want to model that residual spatial autocorrelation, which can often result in getting better spatial predictions and a better fitting model than a non-spatial model. After ESDA, we have to decide 1) what covariates to use, and 2) what spatial covariance functions, Euclidean, tail-up, and tail-down to use, and 3) what shapes of the spatial covariance functions to use. The shapes, or forms, of spatial covariance functions are described using terms such as nugget, exponential, spherical, or Gaussian, etc. Examples of those shapes are shown below.

```{r shapes}
show.vgms()
```

At the Geospatial Data Science in R site <https://zia207.github.io/geospatial-r-github.io/index.html> see the section on spatial interpolation.