---
title: "R_vdeq_toregram"
author: "Michael G. McManus, US EPA/ORD, and Travis Linscome-Hatifield, Neptune and Company"
date: "04/03/2025"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline
For each of the scripts, 01_Semivariogram.Rmd through 06_SSN_Model_Predict.Rmd, begin with a cleared Environment.

Our exploratory spatial data analysis (ESDA) continues by examining spatial variation now over stream network distances, instead of only Euclidean distance. 

## Libraries and Refernces
```{r libraries}
library(tidyverse)
library(gstat) # for semivariograms
library(SSN2) # for spatial stream network (SSN) objects
library(janitor) # clean_names function

sessionInfo()

# attached base packages:
# [1] stats     graphics  grDevices utils     datasets  methods   base     
# 
# other attached packages:
#  [1] janitor_2.2.0   SSN2_0.2.1      gstat_2.1-1     lubridate_1.9.3 forcats_1.0.0  
#  [6] stringr_1.5.1   dplyr_1.1.4     purrr_1.0.2     readr_2.1.5     tidyr_1.3.1    
# [11] tibble_3.2.1    ggplot2_3.5.1   tidyverse_2.0.0



```

## SSN Import & Extract Data Frame

A spatial stream network object is needed to evaluate semivariance in the response variable as a function of Euclidean and stream network distances, flow-connected and flow-unconnected. 

Ideally for an SSN analysis, the data set has these 3 characteristics of monitoring sites:
* Sample size:  minimum of 50, maximum of 2,000
* Spatial Configuration:  spans headwaters to outlet
* Spatial Clustering:  around confluences.

We import the rfbc SSN object and extract the data frame, DFobs, which holds the attribute data and "sticky" geometry. We can manipulate, transform, etc. the DFobs and put it back into the SSN object.
```{r ssn_import}
j_ssn1a <- SSN2::ssn_import("ssn_object/James_071024_pluspreds.ssn", predpts = "sites")
class(j_ssn1a)

names(j_ssn1a)
names(j_ssn1a$obs)

# for consistency I am going to pull the obs from the ssn and apply clean_names so all variable names are in lower case

DFobs <- SSN2::ssn_get_data(j_ssn1a)
DFobs <- clean_names(DFobs)
names(DFobs)

# this shows that DFobs is both an sf and data.frame
class(DFobs)

# note ssn_put_data requires sf object and SSN2 object
# this is putting cleaned names of DFobs back into SSN2 object
j_ssn1a <-  SSN2::ssn_put_data(DFobs,j_ssn1a)

```

## Create Distance Matrix
SSN2 creates a distance folder where distance R object stored. Look where the SSN object is located in File Explorer. In our case we are looking for the James_071024_pluspreds.ssn folder and within that is the distance folder. The creation of the distance matrix to calculate all three distant types only needs to be done once as long as the geography of points (observation or sites, and prediction points) and flowlines (edges) have not been altered. The function has already been run so it is commented out. Note that this distance matrix only contains distances for obs. When we go to make predictions with an SSN model we need to make sure and run the distance matrix and specify that is also to include distances for preds.
```{r distance_matrix}
## Create distance matrices for observed sites
# SSN2::ssn_create_distmat(j_ssn2, overwrite = TRUE)
```

## Torgegram

We are still interested in answering the question:  do sites nearer to each other have smaller semivariances that sites further apart. Now, we are asking that question over Euclidean, flow-unconnected, and flow-connected distances.

### Euclidean Distance
We will begin with gamma, or semivariance, plotted as a function of Euclidean distance so this is just a semivariogram plot. Recall that our individual vsci observations for each station are binned over distance giving an average distance and average gamma based on the number of pairs of points used for those averages. The size of the point plotted in the graph corresponds to the number of pairs of points.
```{r torg_euclid}
names(j_ssn1a)
summary(j_ssn1a)

ztg <- SSN2::Torgegram(vscivcpmi ~ 1, j_ssn1a, type = c( "euclid"))
class(ztg)
plot(ztg, main = "VSCI:  Euclidean")
names(ztg)

View(ztg$euclid)

```

Notice in the View of ztg$euclid that columns names of dist, gamma, and np correspond to the x-axis, y-axis, and points plotted in the graph, respectively. Back in script 01 the variogram function in the gstat package has a default cutoff of 1/3 the bounding box diagonal, which gave a distance cutoff of ~ 100km. In the SSN2 package, the Torgegram function uses a default cutoff of 1/2 the bounding box diagonal, which gives a Euclidean distance cutoff of ~ 150 km.

Besides the default plot based on the Torgegram object, we can also use ggplot. 

```{r ggplot_euclid}
torg_eu <- ztg[["euclid"]]
names(torg_eu)
class(torg_eu)
ggplot(torg_eu, aes(x=dist, y=gamma,size=np)) + geom_point() + ggtitle("VSCI Euclidean Semivariogram")

rm(ztg)
```

### Flow-Unconnected Distance
Stations are flow-unconnected when they are on different branches of the stream network and share a confluence. The calculations of average gamma and average distance remain the same, but are restricted to those pairs of stations that are flow-unconnected.

```{r torg_flow_unconnect}
ztg <- SSN2::Torgegram(vscivcpmi ~ 1, j_ssn1a, type = c( "flowuncon"))

plot(ztg, main = "VSCI:  Flow-Unconnected")

View(ztg$flowuncon)

```

Now using ggplot.
```{r ggplo_flowuncon}
torg_fu <- ztg[["flowuncon"]]

ggplot(torg_fu, aes(x=dist, y=gamma,size=np)) + geom_point() + ggtitle("VSCI Flow-Unconnected Torgegram")

rm(ztg)

```

### Flow-Connected Distance
Stations are flow-connected when they have an upstream to downstream relationship. The calculations of average gamma and average distance remain the same, but are restricted to those pairs of stations that are flow-unconnected. Why might the flow-connected Toregram look so wonky? 
```{r torg_flow_unconnect}
ztg <- SSN2::Torgegram(vscivcpmi ~ 1, j_ssn1a, type = c( "flowcon"))

plot(ztg, main = "VSCI:  Flow-Connected")

View(ztg$flowcon)

```

If you examine the flow-connected data frame created above you will see the incredibly small sample size for the 15 plotted points. One typically wants the number of pairs of points (np) to be at least 30 for each semivariance point. 

Now using ggplot
```{r ggplot_flowcon}
torg_fc <- ztg[["flowcon"]]
names(torg_fc)
class(torg_fc)
ggplot(torg_fc, aes(x=dist, y=gamma,size=np)) + geom_point() + ggtitle("VSCI Flow-Connected Torgegram")

rm(ztg)
```

### Variation of Torgegram Plotting
```{r toregrams-together_separate}
ztg <- SSN2::Torgegram(vscivcpmi ~ 1, j_ssn1a, type = c( "flowcon", "flowuncon", "euclid"))

plot(ztg, main = "VSCI:  All Distances")

# specifying separate = TRUE requires hitting the return bar to see each Torgegram. If trying to run all the code chunks from top to bottom at once, then comment this line out otherwise it will error.

# plot(ztg, separate = TRUE, main = "VSCI")
```

The Euclidean semivariogram suggest some spatial structure at distances up to 75 km apart. The flow-unconnected Torgegram suggests some spatial structure as well. As noted, the flow-connected Torgegram does not look spatially interpretable. How did the manner in which the sites were selected likely affect the wonky flow-connected pattern? 

All of the plots above accepted the defaults for cutoff and equally spaced bins, which is 15.

We have evidence of spatial autocorrelation in VSCI. But, what we want to know is after we have accounted for variation in VSCI by modeling using covariates, is there leftover, or residual, spatial autocorrelation. If that's the case, then we want to model that residual spatial autocorrelation, which can often result in getting better spatial predictions and a better fitting model than a non-spatial model. After ESDA, we have to decide 1) what covariates to use, 2) what spatial covariance functions, Euclidean, tail-up, and tail-down to use, and 3) what shapes of the spatial covariance functions to use. The shapes, or forms, of spatial covariance functions are described using terms such as nugget, exponential, spherical, Gaussian, etc. Examples of those shapes are shown below.

```{r shapes}
show.vgms()
```

During exploratory spatial data analysis we do not try specifying the shapes of the spatial autocovariance functions. That is done during the development of an SSN model.

At the Geospatial Data Science in R site <https://zia207.github.io/geospatial-r-github.io/index.html> see the section on spatial interpolation.