---
title: "ssn_vdeq_tutdeu_comparison"
author: "mcmanus"
date: "12/03/2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

The purpose of this script is to compare 15 spatial autocovariances for a y-intercept only model. The model with the smallest AICc, RMSPE, and highest cor2 will be used in the VDEQ SSN models with covariates. The table of the 15 model comparisons was saved on 12/03/2024. The 	
td_exp_eu_exp model was selected.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(SSN2)
library(tidyverse)

j_ssn1a <- SSN2::ssn_import("ssn_object/James_071024_pluspreds.ssn", predpts = "sites")
class(j_ssn1a)

names(j_ssn1a)
names(j_ssn1a$obs)

# for consistency I am going to pull the obs from the ssn and apply clean_names so all variable names are in lower case

DFobs <- SSN2::ssn_get_data(j_ssn1a)
DFobs <- clean_names(DFobs)
names(DFobs)

# this shows that DFobs is both an sf and data.frame
class(DFobs)

# note ssn_put_data requires sf object and SSN2 object
# this is putting cleaned names of DFobs back into SSN2 object
j_ssn1a <-  SSN2::ssn_put_data(DFobs,j_ssn1a)

```

# 1.0 One Way:  SSN Y-Intercept
All these models have the same fixed effects structure so REML estimation, the default, can be used.
```{r oneway_ssn_yintercept}
tu_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "exponential",
  taildown_type = "none",
  euclid_type = "none",
  additive = "afv_area"
)
summary(tu_exp)
varcomp(tu_exp)

# plot(tu_exp, which = c(1:6))

tu_exp_res <- as.data.frame(tu_exp$residuals)

ggplot(tu_exp_res, aes(x=standardized)) + geom_histogram()
ggplot(tu_exp_res, aes(x=standardized)) + geom_boxplot()
ggplot(tu_exp_res, aes(x=standardized)) + geom_density()
# maximum stream and Euclidean distances
tu_exp$euclid_max # this matches my Euclid max
tu_exp$tail_max # this is much larger than my stream max




tu_lin <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "linear",
  taildown_type = "none",
  euclid_type = "none",
  additive = "afv_area"
)

td_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "none",
  additive = "afv_area"
)

td_lin <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "linear",
  euclid_type = "none",
  additive = "afv_area"
)

eu_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "none",
  euclid_type = "exponential",
  additive = "afv_area"
)

eu_gau <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "none",
  euclid_type = "gaussian",
  additive = "afv_area"
)

models_yintercept <- glances(tu_exp, tu_lin, td_exp, td_lin, eu_exp, eu_gau)
loocv(tu_exp)
loocv(tu_lin)
loocv(td_exp)
loocv(td_lin)
loocv(eu_exp)
loocv(eu_gau)
```

# 2.0 Two Way:  SSN Y-Intercept
```{r twoway_ssn_yintercept}
tu_exp_td_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "exponential",
  taildown_type = "exponential",
  euclid_type = "none",
  additive = "afv_area"
)
summary(tu_exp_td_exp)
varcomp(tu_exp_td_exp)
tu_exp_td_exp_res <- as.data.frame(tu_exp_td_exp$residuals)

ggplot(tu_exp_td_exp_res, aes(x=standardized)) + geom_histogram()
ggplot(tu_exp_td_exp_res, aes(x=standardized)) + geom_boxplot()
ggplot(tu_exp_td_exp_res, aes(x=standardized)) + geom_density()
# maximum stream and Euclidean distances
tu_exp_td_exp$euclid_max # this matches my Euclid max
tu_exp_td_exp$tail_max # this is much larger than my stream max

tu_exp_td_lin <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "exponential",
  taildown_type = "linear",
  euclid_type = "none",
  additive = "afv_area"
)

tu_lin_td_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "linear",
  taildown_type = "exponential",
  euclid_type = "none",
  additive = "afv_area"
)

tu_lin_td_lin <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "linear",
  taildown_type = "linear",
  euclid_type = "none",
  additive = "afv_area"
)

td_exp_eu_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  additive = "afv_area"
)

summary(td_exp_eu_exp)
varcomp(td_exp_eu_exp)

td_exp_eu_gau <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "gaussian",
  additive = "afv_area"
)


td_lin_eu_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "linear",
  euclid_type = "exponential",
  additive = "afv_area"
)

td_lin_eu_gau <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "none",
  taildown_type = "linear",
  euclid_type = "gaussian",
  additive = "afv_area"
)

models_yintercept <- glances(tu_exp, tu_lin, td_exp, td_lin, eu_exp, eu_gau, tu_exp_td_exp, tu_exp_td_lin, tu_lin_td_exp, tu_lin_td_lin, td_exp_eu_exp, td_exp_eu_gau, td_lin_eu_exp, td_lin_eu_gau)

loocv(tu_exp_td_exp)
loocv(tu_exp_td_lin)
loocv(tu_lin_td_exp)
loocv(tu_lin_td_lin)
loocv(td_exp_eu_exp)
loocv(td_exp_eu_gau)
loocv(td_lin_eu_exp)
loocv(td_lin_eu_gau)

```

# 3.0 Threeway:  SSN Y-Intercept
```{r threeway_ssn_yintercept}
tu_td_eu_exp <- ssn_lm(
  formula = vscivcpmi ~ 1,
  ssn.object = j_ssn1a,
  tailup_type = "exponential",
  taildown_type = "exponential",
  euclid_type = "exponential",
  additive = "afv_area"
)
summary(tu_td_eu_exp)
varcomp(tu_td_eu_exp)
tu_td_eu_exp_res <- as.data.frame(tu_exp_td_exp$residuals)

ggplot(tu_td_eu_exp_res, aes(x=standardized)) + geom_histogram()
ggplot(tu_td_eu_exp_res, aes(x=standardized)) + geom_boxplot()
ggplot(tu_td_eu_exp_res, aes(x=standardized)) + geom_density()
# maximum stream and Euclidean distances
tu_td_eu_exp$euclid_max # this matches my Euclid max
tu_td_eu_exp$tail_max # this is much larger than my stream max

models_yintercept <- glances(tu_exp, tu_lin, td_exp, td_lin, eu_exp, eu_gau, tu_exp_td_exp, tu_exp_td_lin, tu_lin_td_exp, tu_lin_td_lin, td_exp_eu_exp, td_exp_eu_gau, td_lin_eu_exp, td_lin_eu_gau, tu_td_eu_exp)

loocv(tu_td_eu_exp)

write.csv(models_yintercept, file="outputs/models_yintercept.csv",sep=",",row.names=F)
```
