---
title:  "James MLR and SSN Modeling"
author: "Michael G. McManus, US EPA/ORD & Travis Linscome-Hatfield, Neptune and Company"
date: "04/03/2025"
output:
  html_document:
    fig_caption: yes
    number_sections: false
    toc: yes
    toc_float:  yes
    code_folding: hide
    self_contained: yes
    theme: lumen
editor_options: 
  chunk_output_type: console
---

# Outline
The purpose of this script is to model the relationship between VSCI and covariates across the 4 geographies of Watershed (Ws), Watershed-Riparian (WsRp), Catchment (Cat), and Catchment-Riparian (CatRp). For each of the 4 geographies, the modeling is done as follows:

* Use bestglm to select covariates 
* Check model diagnostics of the multiple linear regression (MLR) model
* Run a SSN nugget model (SSN0), which is the equivalent to an MLR model, and check diagnostic plots
* Run the SSN1 with covariates and spatial autocovariances. Note the spatial autocovariances are identified by running script 03_Y_Intercept_SSN_Model.
* Compare SSN0 to SSN1 to see if modeling spatial autocorrelation of residuals improved model fit and model prediction
* Plot loocv observed and predicted
* Map standardized residuals
* Test for random effects of trend stations versus status stations. Trends stations are sampled repeated, with station_id_2 being the identifier that distinguishes trend from status stations. Status stations are only sampled one time.
* Compare SSN models across the 4 geographies using ML estimation method as the models have different covariates.

Portions of this analysis are summarized in the Presentation folder, specifically 2024_sfc_philadelphia_bioassessment_session_mcmanus.pdf. That presentation provides an overview of the detailed analysis presented below.

# Library
```{r setup, collapse=TRUE}

library(tidyverse)
library(janitor)
library(sf)
library(dummy)
library(bestglm)
library(performance)
library(car)
library(SSN2)
library(mapview)
library(leafpop) # for popups in mapview
library(leafsync) # to sync obs and prediction maps

sessionInfo()

# attached base packages:
# [1] stats     graphics  grDevices utils     datasets  methods  
# [7] base     
# 
# other attached packages:
#  [1] leafsync_0.1.0     leafpop_0.1.0      mapview_2.11.2    
#  [4] SSN2_0.2.1         car_3.1-2          carData_3.0-5     
#  [7] performance_0.12.2 see_0.8.5          bestglm_0.37.3    
# [10] leaps_3.1          dummy_0.1.3        sf_1.0-16         
# [13] janitor_2.2.0      lubridate_1.9.3    forcats_1.0.0     
# [16] stringr_1.5.1      dplyr_1.1.4        purrr_1.0.2       
# [19] readr_2.1.5        tidyr_1.3.1        tibble_3.2.1      
# [22] ggplot2_3.5.1      tidyverse_2.0.0  

knitr::opts_chunk$set(message=FALSE, warning=FALSE,collapse = T)

```

# 1.0 Load SSN
```{r ssn_from_script_03}
load("outputs/j_ssn3.RData")

DFobs <- SSN2::ssn_get_data(j_ssn3) %>% clean_names(.)
names(DFobs)
```


# 2.0 Dummy Coding for Bestglm Models
The bestglm function only accepts dummy coding for factors.

# 2.1 Dummy Coding of Factors
Dummy coding for ecoregions, Virginia hydrologic unit subbasins, wetlands, impervious surfaces, and total phosphorus. See the description of dummy coding at <https://stats.oarc.ucla.edu/r/library/r-library-contrast-coding-systems-for-categorical-variables/>.
```{r dummy_code}

# remove geometry as bestglm does not want it
DFobsz <- st_set_geometry(DFobs, NULL)
#creating alternate dummy version for base case comparison if desired
DFobsz_alternate_dummy = DFobsz

# dummy code 5 ecoregions with base being Blue Ridge Mtns
eco_r <- dplyr::select(DFobsz, eco_region)
summary(eco_r)
glimpse(eco_r)

eco_d <- (data.frame(dummy(eco_r)))
class(eco_d)
dim(eco_d)
head(eco_d)
str(eco_d)
distinct(eco_d)
# 5 levels need only n-1 =4 dummy variables, removed base level of Blue Ridge Mountains by dropping first column
eco_d <- eco_d[c(-1)]
dim(eco_d)
head(DFobsz$eco_region)
distinct(eco_d)
head(eco_d)
str(eco_d)
class(eco_d)

DFobsz <- cbind(DFobsz,eco_d)
names(DFobsz)

# dummy code 5 vahusb with base being JU, James Upper
vahusb <- dplyr::select(DFobsz, vahusb)
summary(vahusb)
glimpse(vahusb)

vahusb_d <- (data.frame(dummy(vahusb)))
# 5 levels need only n-1 =4 dummy variables, removed base level of JU by dropping first column
vahusb_d <- vahusb_d[c(-1)]
dim(vahusb_d)
head(DFobsz$vahusb)
distinct(vahusb_d)
head(vahusb_d)
str(vahusb_d)
class(vahusb_d)

DFobsz <- cbind(DFobsz,vahusb_d)
names(DFobsz)


# dummy code 4 watershed wetland levels with base being none
wet_w_f <- dplyr::select(DFobsz, pct_wet_w_f)
summary(wet_w_f)
glimpse(wet_w_f)

wet_w_d <- (data.frame(dummy(wet_w_f)))
# 4 levels need only n-1 =3 dummy variables, removed base level of none by dropping first column
wet_w_d <- wet_w_d[c(-1)]
DFobsz <- cbind(DFobsz,wet_w_d)
names(DFobsz)

# dummy code 4 watershed-riparian wetland levels with base being none
wet_wr_f <- dplyr::select(DFobsz, pct_wet_wr_f)
summary(wet_wr_f)
glimpse(wet_wr_f)

wet_wr_d <- (data.frame(dummy(wet_wr_f)))
# 4 levels need only n-1 =3 dummy variables, removed base level of none by dropping first column
wet_wr_d <- wet_wr_d[c(-1)]
DFobsz <- cbind(DFobsz,wet_wr_d)
names(DFobsz)

# dummy code 4 catchment wetland levels with base being none
wet_c_f <- dplyr::select(DFobsz, pct_wet_c_f)
summary(wet_c_f)
glimpse(wet_c_f)

wet_c_d <- (data.frame(dummy(wet_c_f)))
# 4 levels need only n-1 =3 dummy variables, removed base level of none by dropping first column
wet_c_d <- wet_c_d[c(-1)]
DFobsz <- cbind(DFobsz,wet_c_d)
names(DFobsz)

# dummy code 3 catchment-riparian wetland levels (none, low, high) with base being none
wet_cr_f <- dplyr::select(DFobsz, pct_wet_cr_f)
summary(wet_cr_f)
glimpse(wet_cr_f)

wet_cr_d <- (data.frame(dummy(wet_cr_f)))
# 3 levels need only n-1 =2 dummy variables, removed base level of none by dropping first column
wet_cr_d <- wet_cr_d[c(-1)]
DFobsz <- cbind(DFobsz,wet_cr_d)
names(DFobsz)

# dummy code 3 catchment-riparian impervious surfaces levels with base being none
imp_cr_f <- dplyr::select(DFobsz, pct_imp_rp_c_f)
summary(imp_cr_f)
glimpse(imp_cr_f)

imp_cr_d <- (data.frame(dummy(imp_cr_f)))
# 3 levels need only n-1 =2 dummy variables, removed base level of none by dropping first column
imp_cr_d <- imp_cr_d[c(-1)]
DFobsz <- cbind(DFobsz,imp_cr_d)
names(DFobsz)

# dummy code 3 total phosphorus levels with base being low
tp_f <- dplyr::select(DFobsz, tp_f)
summary(tp_f)
glimpse(tp_f)

tp_d <- (data.frame(dummy(tp_f)))
# 3 levels need only n-1 =2 dummy variables, removed base level of none by dropping first column
tp_d <- tp_d[c(-1)]
DFobsz <- cbind(DFobsz,tp_d)
names(DFobsz)

# remove files not needed downstream
rm(eco_d, eco_r, imp_cr_d, imp_cr_f, tp_d, tp_f, vahusb, vahusb_d, wet_c_d, wet_c_f, wet_cr_d, wet_cr_f, wet_w_d, wet_w_f, wet_wr_d, wet_wr_f)
```

# 2.15 Dummy Coded to SF to SSN
The 23 dummy coded variables just created need to be put into an SF object and then, later, into an SSN object. 
```{r dummy_sf_ssn}
names(DFobsz)

DFobsz2 <- dplyr::select(DFobsz, c(station_id_2, eco_region_Central.Appalachian.Ridges.and.Valleys:tp_f_high))

# put dummy covariates in an SF object
DFobs2 <- full_join(DFobs, DFobsz2, by = join_by(station_id_2))
class(DFobs2)
names(DFobs2)


```

# 3.0 SSN Ws, WsRp, Cat, & CatRp Models
Before run check_model function below have Plots tab fully expanded vertically. The check_model functions provides visual check of various model assumptions. The added variable plots, also called partial regression plots, are made using the avPlots function from the car R package and give a visualization of the slope estimates.

The candidate pool of 21 covariates came from these categories:  1) climate 2) landcover, 3) instream measurements, and 4) geophysical or hydrologic units.

The maximum number of covariates to be retained during bestglm selection was set to 4 before any modeling was done. This was done to produce a conservative model based on the number of observations relative to the number of parameters estimated in the model. Four slopes and 1 y-intercept are estimated with nvmax = 4. Then, a spatial stream network model could include up to 7 additional parameters, with 2 parameters estimated each for Euclidean, Tail Up, and Tail down autocovariances plus an independent error, or nugget, parameter. For 199 observations, there could be 12 model parameters or ~ 17 observations/parameter, and that seemed a good way to avoid overfitting the model as a rule of thumb is to have 10 observations/parameter.

# 4.0 Ws-Wq
Analysis done at watershed extent.

### 4.1 Bestglm and SSN0 Nugget Model
The X1 matrix below includes the dummy factors with the station_id_2 enables the join the dummy factor to an sf object and then to the SSN. 

```{r ws_bestglm}

# candidate covariates
WsWq <- DFobsz|>
  dplyr::select(precip_mm, tmean, pct_imp_w_emplog, pct_for_w_emplog, pct_hay_w_emplog, pct_grs_w_emplog, elev_ws, do, p_h, tothab, l_spc, l_tds, l_turb, l_tn, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, bin_wet_w, tp_f_medium, tp_f_high)
str(WsWq)
# X can only contain numeric or factor
WsWq<- WsWq %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL','tp_f_medium', 'tp_f_high'), as.numeric)

# bind matrix of covariates to response variable
WsWqy<-cbind.data.frame(WsWq,y=DFobsz$vsci)
ASRcum_wswq1.best<-bestglm(WsWqy,IC = "AIC", nvmax =4, TopModels=15)

# See what terms are in best models
ASRcum_wswq1.best$BestModels
ASRcum_wswq1.best$Subsets
summary(ASRcum_wswq1.best$BestModel)

# create mlr model so diagnostics can be run on that lm object
wswq_mlry2 <- lm(vsci ~ pct_imp_w_emplog + elev_ws + do + vahusb_JL, data = DFobsz)
summary(wswq_mlry2)
check_model(wswq_mlry2)
# if you wish to view avPlots uncomment below. But this hangs if you attempt to run the whole script as one block
# avPlots(wswq_mlry2)

# SSN0 nugget model
# station_id_2 lets join_by work
X1 <- DFobsz|>
   dplyr::select(station_id_2, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, pct_wet_w_f_low, pct_wet_w_f_medium, pct_wet_w_f_high, tp_f_medium, tp_f_high)

# X can only contain numeric or factor
X1 <- X1 %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL', 'pct_wet_w_f_low', 'pct_wet_w_f_medium', 'pct_wet_w_f_high', 'tp_f_medium', 'tp_f_high'), as.numeric)
str(X1)
# put dummy covariates into SF object by full join to original DFobs
DFobs3a <- full_join(DFobs, X1, by = join_by(station_id_2))
names(DFobs3a)

# put SF object into SSN
j_ssn3 <-  SSN2::ssn_put_data(DFobs3a,j_ssn3)

# ssn0 Watershed
ssn0_wswq_mlr <- ssn_lm(
  formula = vsci ~ pct_imp_w_emplog + elev_ws + do + vahusb_JL,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "none",
  euclid_type = "none",
  nugget_type = "nugget",
  estmethod = "reml" #default
)
summary(ssn0_wswq_mlr)
varcomp(ssn0_wswq_mlr)
loocv(ssn0_wswq_mlr)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn0_wswq_mlr, which = c(1:6))
```

### 4.2 SSN0 Nugget vs SSN1
This chunk compares the nugget model, typical nonspatial MLR or SSN0, to the SSN1 model that specifies the spatial autocovariances that were identified in script 03_Y-Intercept_SSN_Model. The comparison is made using model fit diagnostics, AICc, and leave-one-out-cross-validation (LOOCV) metrics of the root mean squared prediction error (RMSPE) and the squared correlation between the observed value and its leave-one-out prediction (cor2).

The models_yintercept.csv is from script 03 is in the outputs folder.
```{r nugget_vs_spatial_ws}
ssn_wswq_reml1 <- ssn_lm(
  formula = vsci ~ pct_imp_w_emplog + elev_ws + do + vahusb_JL,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area"
)
summary(ssn_wswq_reml1)
varcomp(ssn_wswq_reml1)
loocv(ssn_wswq_reml1)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_wswq_reml1, which = c(1:6))

models_ssn0_ssn1ws <- glances(ssn0_wswq_mlr, ssn_wswq_reml1)
print(models_ssn0_ssn1ws)


```


## Synced map of Obs, Fitted, Standardized Residuals
In a spatial model (which includes spatial stream network models), residuals are spatially autocorrelated, so it is more helpful to apply typical model diagnostics to the standardized residuals, which have been decorrelated. The map below of standardized residuals can give a spatial characterization. Are extreme values of the standardized residuals spread throughout the spatial domain, or is there a geographic clustering of extreme standardized residuals? A clustering might indicate a locality that other covariates should be considered.

```{r synced_maps}
# 
# mapview(aug_ssn_wswq_reml1, zcol = ".std.resid", cex = ".std.resid", alpha.regions = .8, legend = TRUE, layer.name = 'Standardized Residuals', popup = popupTable(aug_ssn_wswq_reml1, zcol = c("vsci", ".fitted", ".resid", ".std.resid")))
# 
# # observed vsci at 199 obs
# obs1 <- mapview(aug_ssn_wswq_reml1, zcol = "vsci", cex = "vsci", alpha.regions = .8, legend = TRUE, layer.name = 'VSCI', popup = popupTable(aug_ssn_wswq_reml1, zcol = c("vsci", ".fitted", ".resid", ".std.resid"))) 
# 
# # fitted vsci at 199 obs
# fit1 <- mapview(aug_ssn_wswq_reml1, zcol = ".fitted", cex = ".fitted", alpha.regions = .8, legend = TRUE, layer.name = 'Fitted', popup = popupTable(aug_ssn_wswq_reml1, zcol = c("vsci", ".fitted", ".resid", ".std.resid")))
# 
# # standardized residuals at 199 obs
# stdres1 <- mapview(aug_ssn_wswq_reml1, zcol = ".std.resid", cex = ".std.resid", alpha.regions = .8, legend = TRUE, layer.name = 'Std. Residual', popup = popupTable(aug_ssn_wswq_reml1, zcol = c("vsci", ".fitted", ".resid", ".std.resid")))
# 
# sync(obs1,fit1, ncol=1) # obs & predicted
# sync(fit1, stdres1, ncol=1)
# sync(obs1, fit1, stdres1, ncol=1)
```

### 4.3 Random Effects
The random effect for station_id is coded so the model recognizes that there are 8 stations with repeated 4 to 5 observations whereas all the other stations have single observation. Based on AICc and cor2, ssn_wswq_reml1 is still a better model. 
```{r random_effect_station_id}
ssn_wswq_rand1 <- ssn_lm(
  formula = vsci ~ pct_imp_w_emplog + elev_ws + do + vahusb_JL,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area",
  random = ~ (1 | station_id)
)

summary(ssn_wswq_rand1)
varcomp(ssn_wswq_rand1)
loocv(ssn_wswq_rand1)

# if want to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key
# plot(ssn_wswq_rand1, which = c(1:6))

models_rand1 <- glances(ssn0_wswq_mlr, ssn_wswq_reml1, ssn_wswq_rand1)
print(models_rand1)

```

# 5.0 WsRp-Wq
Analysis done at Watershed-Riparian extent.

### 5.1 Bestglm and SSN0 Nugget Model

```{r wsriparian_bestglm}
WsRpWq <- DFobsz|>
  dplyr::select(precip_mm, tmean, pct_imp_rp_w_emplog, pct_for_wr_emplog, pct_hay_wr_emplog, pct_grs_wr_emplog, elev_ws, do, p_h, tothab, l_spc, l_tds, l_turb, l_tn,  vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, bin_wet_wr, tp_f_medium, tp_f_high)
str(WsRpWq)
# X can only contain numeric or factor
WsRpWq<- WsRpWq %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL', 'tp_f_medium', 'tp_f_high'), as.numeric)

# bind matrix of covariates to response variable
WsRpWqy<-cbind.data.frame(WsRpWq,y=DFobsz$vsci)
ASRcum_wsrpwq1.best<-bestglm(WsRpWqy,IC = "AIC",nvmax =4, TopModels=15)

# See what terms are in best models
ASRcum_wsrpwq1.best$BestModels
ASRcum_wsrpwq1.best$Subsets
summary(ASRcum_wsrpwq1.best$BestModel)

wsrpwq_mlry2 <- lm(vsci ~ pct_imp_rp_w_emplog + elev_ws + do + tothab, data = DFobsz)
summary(wsrpwq_mlry2)
check_model(wsrpwq_mlry2)
# if you wish to view avPlots uncomment below. But this hangs if you attempt to run the whole script as one block
# avPlots(wsrpwq_mlry2)

# SSN0 nugget model
# station_id_2 lets join_by work
X2 <- DFobsz|>
  dplyr::select(station_id_2, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, pct_wet_wr_f_low, pct_wet_wr_f_medium, pct_wet_wr_f_high, tp_f_medium, tp_f_high)

# X can only contain numeric or factor
X2 <- X2 %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL', 'pct_wet_wr_f_low', 'pct_wet_wr_f_medium', 'pct_wet_wr_f_high', 'tp_f_medium', 'tp_f_high'), as.numeric)

# put dummy covariates in an SF object
DFobs3a <- full_join(DFobs, X2, by = join_by(station_id_2))
names(DFobs3a)

# put SF object into SSN
j_ssn3 <-  SSN2::ssn_put_data(DFobs3a,j_ssn3)

# SSN0 Watershed-Riparian
ssn0_wsrpwq_mlr <- ssn_lm(
  formula = vsci ~ pct_imp_rp_w_emplog + elev_ws + do + tothab,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "none",
  euclid_type = "none",
  nugget_type = "nugget",
  estmethod = "reml"
)
summary(ssn0_wsrpwq_mlr)
varcomp(ssn0_wsrpwq_mlr)
loocv(ssn0_wsrpwq_mlr)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn0_wsrpwq_mlr, which = c(1:6))
```

### 5.2 SSN0 Nugget vs SSN1
```{r nugget_vs_spatial_wsriparian}
ssn_wsrpwq_reml1 <- ssn_lm(
  formula = vsci ~ pct_imp_rp_w_emplog + elev_ws + do + tothab,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area"
)
summary(ssn_wsrpwq_reml1)
varcomp(ssn_wsrpwq_reml1)
loocv(ssn_wsrpwq_reml1)


#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_wsrpwq_reml1, which = c(1:6))

models_ssn0_ssn1wsrp <- glances(ssn0_wsrpwq_mlr, ssn_wsrpwq_reml1)
print(models_ssn0_ssn1wsrp)

```


### 5.3 Random Effects: Ws_Riparian
```{r wsrpwq_random}
ssn_wsrpwq_rand1 <- ssn_lm(
  formula = vsci ~ pct_imp_rp_w_emplog + elev_ws + do + tothab,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area",
  random = ~ (1 | station_id)
)

summary(ssn_wsrpwq_rand1)
varcomp(ssn_wsrpwq_rand1)
loocv(ssn_wsrpwq_rand1)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_wsrpwq_rand1, which = c(1:6))

models_rand2 <- glances(ssn0_wsrpwq_mlr, ssn_wsrpwq_reml1, ssn_wsrpwq_rand1)
print(models_rand2)

```

### ssn_wsrpwq_reml1 Mapview residuals 
Standardized residuals can be helpful as described previously.

```{r maps_ssn_wsrpwq_reml1}
mapview(aug_ssn_wsrpwq_reml1, zcol = ".std.resid",  alpha.regions = .8, legend = TRUE, layer.name = 'Standardized Residuals', popup = popupTable(aug_ssn_wsrpwq_reml1, zcol = c("vsci", ".fitted", ".resid", ".std.resid")))
```

# 6.0 Cat-Wq
Analysis done at Catchment extent.

```{r cat_bestglm}
CatWq <- DFobsz|>
  dplyr::select(precip_mm, tmean, pct_imp_c_emplog, pct_for_c_emplog, pct_hay_c_emplog, bin_grs_c, elev_cat, do, p_h, tothab, l_spc, l_tds, l_turb, l_tn, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, bin_wet_c,  tp_f_medium, tp_f_high)
str(CatWq)
# X can only contain numeric or factor
CatWq<- CatWq %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL', 'bin_wet_c','tp_f_medium', 'tp_f_high'), as.numeric)

# bind matrix of covariates to response variable
CatWqy<-cbind.data.frame(CatWq,y=DFobsz$vsci)
ASRcum_catwq1.best<-bestglm(CatWqy,IC = "AIC",nvmax =4, TopModels=15)

# See what terms are in best models
ASRcum_catwq1.best$BestModels
ASRcum_catwq1.best$Subsets
summary(ASRcum_catwq1.best$BestModel)

catwq_mlry2 <- lm(vsci ~ pct_imp_c_emplog + elev_cat + tothab + l_tn, data = DFobsz)
summary(catwq_mlry2)
check_model(catwq_mlry2)
# if you wish to view avPlots uncomment below. But this hangs if you attempt to run the whole script as one block
# avPlots(catwq_mlry2)

# No categorical covariates needed so commented out X3
# SSN0 nugget model
# st_id_tren lets join_by work
# X3 <- DFobsz|>
#  dplyr::select(st_id_tren, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, pct_wet_c_f_low, pct_wet_c_f_medium, pct_wet_c_f_high, tp_f_medium, tp_f_high)

# X can only contain numeric or factor
# X3 <- X3 %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL', 'pct_wet_c_f_low', 'pct_wet_c_f_medium', 'pct_wet_c_f_high','tp_f_medium', 'tp_f_high'), as.numeric)

# put dummy covariates in an SF object
# DFobs3a <- full_join(DFobs2a, X3, by = join_by(st_id_tren))
# names(DFobs3a)
# put transformed covariates in an SF object
# DFobs2a <- full_join(DFobs2, DFobsz2, by = join_by(st_id_tren))

# put SF object into SSN
# j_ssn3 <-  SSN2::ssn_put_data(DFobs3a,j_ssn3)

# SSN0 Catchment
ssn0_catwq_mlr <- ssn_lm(
  formula = vsci ~ pct_imp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "none",
  euclid_type = "none",
  nugget_type = "nugget",
  estmethod = "reml" #default
)
summary(ssn0_catwq_mlr)
varcomp(ssn0_catwq_mlr)
loocv(ssn0_catwq_mlr)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn0_catwq_mlr, which = c(1:6))
```

### 6.2 SSN0 Nugget vs SSN1
```{r nugget_vs_spatial_cat}
ssn_catwq_reml1 <- ssn_lm(
  formula = vsci ~ pct_imp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area"
)
summary(ssn_catwq_reml1)
varcomp(ssn_catwq_reml1)
loocv(ssn_catwq_reml1)


#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_catwq_reml1, which = c(1:6))

models_ssn0_ssn1cat <- glances(ssn0_catwq_mlr, ssn_catwq_reml1)
print(models_ssn0_ssn1cat)

aug_ssn_catwq_reml1 <- augment(ssn_catwq_reml1, drop = FALSE)
class(aug_ssn_catwq_reml1)
mapview(aug_ssn_catwq_reml1$.std.resid)

```

### 6.3 Random Effects: Cat-Wq
Check random effects of station_id on ssn_catwq_rand1.
```{r catwq_random}
ssn_catwq_rand1 <- ssn_lm(
  formula = vsci ~ pct_imp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area",
  random = ~ (1 | station_id)
)
summary(ssn_catwq_rand1)
varcomp(ssn_catwq_rand1)
loocv(ssn_catwq_rand1)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_catwq_rand1, which = c(1:6))

models_rand3 <- glances(ssn0_catwq_mlr, ssn_catwq_reml1,ssn_catwq_rand1)
print(models_rand3)
```

# 7.0 CatRp-Wq
Analysis done at Catchment-Riparian extent.

With this smallest geography (by extent and configuration), I have to shift some land cover from continuous to categorical. The previous wetland factor with multiple levels was replaced with a binary absence/presence variable (bin_wet_cr) as was grass cover. Grass at this geography has a median value of 0 so I made an absence/presence variable for grass cover (bin_grs_cr) so a form of that covariate is in the candidate pool of covariates.

### 7.1 Bestglm and SSN0 Nugget Model
```{r catriparian_bestglm}
CatRpWq <- DFobsz|>
  dplyr::select(precip_mm, tmean, pct_imp_rp_c_emplog, pct_for_cr_emplog, pct_hay_cr_emplog, bin_grs_cr, elev_cat, do, p_h, tothab, l_spc, l_tds, l_turb, l_tn,  vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, bin_wet_cr, tp_f_medium, tp_f_high)
str(CatRpWq)
# X can only contain numeric or factor
CatRpWq<- CatRpWq %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'vahusb_JA', 'vahusb_JL',
'tp_f_medium', 'tp_f_high'), as.numeric)

# bind matrix of covariates to response variable
CatRpWqy<-cbind.data.frame(CatRpWq,y=DFobsz$vsci)
ASRcum_catrpwq1.best<-bestglm(CatRpWqy,IC = "AIC",nvmax =4, TopModels=15)

# See what terms are in best models
ASRcum_catrpwq1.best$BestModels
ASRcum_catrpwq1.best$Subsets
summary(ASRcum_catrpwq1.best$BestModel)

catrpwq_mlry2 <- lm(vsci ~ pct_imp_rp_c_emplog + elev_cat + tothab + l_tn, data = DFobsz)
summary(catrpwq_mlry2)
check_model(catrpwq_mlry2)
# if you wish to view avPlots uncomment below. But this hangs if you attempt to run the whole script as one block
# avPlots(catrpwq_mlry2)

# No categorical predictors selected so comment out code
# SSN0 nugget model
# st_id_tren lets join_by work
# X4 <- DFobsz|>
#   dplyr::select(st_id_tren, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, pct_wet_cr_f_low, pct_wet_cr_f_high, tp_f_medium, tp_f_high)

# X can only contain numeric or factor
# X4 <- X4 %>% mutate_at(c('vahusb_JM', 'vahusb_JR', 'eco_region_Piedmont', 'vahusb_JA', 'pct_wet_cr_f_low', 'pct_wet_cr_f_high', 'tp_f_medium', 'tp_f_high'), as.numeric)

# put dummy covariates in an SF object
# DFobs3a <- full_join(DFobs2a, X4, by = join_by(st_id_tren))
# names(DFobs3a)
# put transformed covariates in an SF object
# DFobs2a <- full_join(DFobs2, DFobsz2, by = join_by(st_id_tren))

# put SF object into SSN
# j_ssn3 <-  SSN2::ssn_put_data(DFobs3a,j_ssn3)

# SSN0 Catchment
ssn0_catrpwq_mlr <- ssn_lm(
  formula = vsci ~ pct_imp_rp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "none",
  euclid_type = "none",
  nugget_type = "nugget",
  estmethod = "reml" #default
)
summary(ssn0_catrpwq_mlr)
varcomp(ssn0_catrpwq_mlr)
loocv(ssn0_catrpwq_mlr)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn0_catrpwq_mlr, which = c(1:6))

```

### 7.2 SSN0 vs SSN1
```{r nugget_vs_spatial_catriparian}
ssn_catrpwq_reml1 <- ssn_lm(
  formula = vsci ~ pct_imp_rp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area"
)
summary(ssn_catrpwq_reml1)
varcomp(ssn_catrpwq_reml1)
loocv(ssn_catrpwq_reml1)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_catrpwq_reml1, which = c(1:6))

models_ssn0_ssn1catrp <- glances(ssn0_catrpwq_mlr, ssn_catrpwq_reml1)
print(models_ssn0_ssn1catrp)

aug_ssn_catrpwq_reml1 <- augment(ssn_catrpwq_reml1, drop = FALSE)

# mapview(aug_ssn_catrpwq_reml1$.std.resid)

```

### 7.3 Random Effects: CatRp-Wq
```{r catrpwq_random}
ssn_catrpwq_rand1 <- ssn_lm(
  formula = vsci ~ pct_imp_rp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "reml",
  additive = "afv_area",
  random = ~ (1 | station_id)
)
summary(ssn_catrpwq_rand1)
varcomp(ssn_catrpwq_rand1)
loocv(ssn_catrpwq_rand1)

#to run script continuously from top to bottom comment out plot statement below as requires repeated hitting of enter key, which skips script lines in doing so
# plot(ssn_catrpwq_reml1, which = c(1:6))

models_rand4 <- glances(ssn0_catrpwq_mlr, ssn_catrpwq_reml1, ssn_catrpwq_rand1)
print(models_rand4)
```

# 8.0 Compare Four Geography SSN Models with ML
This to compare different SSN models from the 4 geographies. Given that the models have different covariates then the estimation method needs to be maximum likelihood or ML for AICc comparisons. See An Introduction to Spatial Stream Network Modeling in R using SSN2 for a description of when REML vs ML is used. That discussion is the paragraph just above section 4.1.1 Model Summaries. Note to evaluate the four models in predictive performance the loocv function is used and comparisons of RMSPE and cor2 are made.

The watershed model, ssn_wswq_ml1, fits the data better and predicts better than the 3 other models. We will now use the watershed model for predictions in script 06.
```{r ssn_ml}
ssn_wswq_ml1 <- ssn_lm(
  formula = vsci ~ pct_imp_w_emplog + elev_ws + do + vahusb_JL,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "ml",
  additive = "afv_area"
)

ssn_wsrpwq_ml1 <- ssn_lm(
  formula = vsci ~ pct_imp_rp_w_emplog + elev_ws + do + tothab,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "ml",
  additive = "afv_area"
)

ssn_catwq_ml1 <- ssn_lm(
  formula = vsci ~ pct_imp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "ml",
  additive = "afv_area"
)

ssn_catrpwq_ml1 <- ssn_lm(
  formula = vsci ~ pct_imp_rp_c_emplog + elev_cat + tothab + l_tn,
  ssn.object = j_ssn3,
  tailup_type = "none",
  taildown_type = "exponential",
  euclid_type = "exponential",
  nugget_type = "nugget",
  estmethod = "ml",
  additive = "afv_area"
)

models_ml_ssn <- glances(ssn_wswq_ml1, ssn_wsrpwq_ml1, ssn_catwq_ml1, ssn_catrpwq_ml1)

print(models_ml_ssn)

loocv(ssn_wswq_ml1)
loocv(ssn_wsrpwq_ml1)
loocv(ssn_catwq_ml1)
loocv(ssn_catrpwq_ml1)


```

# Neptune Code from TLH
This code is included for thoroughness, but does not need to be run unless user wants to see output.

In addition to reviewing R code, Neptune statistician Travis Linscome-Hatfield provided code chunks that evaluated 1) assigning a different reference group, rather than the default, and 2) did having a different reference group result in a different model being selected by the bestglm function. Running these two chunks resulted in the same model being selected based on the default group of JU.

## TLH Chunk 1 Dummy Coding of Factors-Travis
Travis Linscome-Hatfield at Neptune provide the code chunk below to see if changing the reference group altered the analysis results and conclusions. This evaluation was motivated by the observation that the default reference group for the Virginia hydrologic unit subbasins was JU, the upper James.
Dummy coding for ecoregions, Virginia hydrologic unit subbasins, wetlands, impervious surfaces, and total phosphorus.
```{r dummy_code-Travis}

names(DFobsz_alternate_dummy)
# the 3 commented lines below simply allow for variable definition for quick funciton adjustment.
dat = DFobsz_alternate_dummy
alt_dat = DFobsz_alternate_dummy
variable = "eco_region"
category_break = function(dat,alt_dat,variable,base1=NULL,base2=NULL){
  
  #pull the column we wish to dummy code
  var_r <- dplyr::select(dat, all_of(variable))
  # get the value with the highest number of instances
  var_b <- as.data.frame(table(var_r))
  var_b <- var_b[order(-var_b$Freq),]
  # if base cases aren't already defined
  if(is.null(base1)){
    # pick base case
    base1 = as.character(var_b[1,1])
  }
  
  # if base2 isn't defined 
  if(is.null(base2)){
    if(base1 == as.character(var_b[1,1])){
      base2 = as.character(var_b[2,1])
    }else{
      base2 = as.character(var_b[1,1])
    }
  }

  
  # dummying the variable
  var_d <- (data.frame(dummy(var_r)))
  # identify which column is the base case
  # turning column names into something that can be directly matched with the base case
  col_short = colnames(var_d)
  col_short = gsub(variable,"",col_short)
  col_short = gsub("_","",col_short)
  col_short = gsub("\\."," ",col_short)
  base1_col_id = which(base1 == col_short)
  base2_col_id = which(base2 == col_short)

  
  # remove the base case from var_d
  var_d1 = var_d[,-base1_col_id]
  var_d2 = var_d[,-base2_col_id]
  # binding into apropriate datasets
  dat = cbind(dat,var_d1)
  alt_dat = cbind(alt_dat,var_d2)
  
  out_list = list("Variable"=variable,
                  "base1"=base1,
                  "base2"=base2,
                  "dat"=dat,
                  "alt_dat"=alt_dat)
  return(out_list)
  
}

# dummy data for eco_region
df_dum_eco = category_break(dat= DFobsz_alternate_dummy,
                            alt_dat=DFobsz_alternate_dummy,
                            "eco_region")
# dummy data for vahusb
df_dum_vahusb = category_break(df_dum_eco$dat,
                               df_dum_eco$alt_dat,
                               "vahusb")
# dummy data for wetland (pct_wet_w_f)
levels(DFobsz_alternate_dummy$pct_wet_w_f)
df_dum_wet = category_break(dat = df_dum_vahusb$dat,
                            alt_dat = df_dum_vahusb$alt_dat,
                            variable = "pct_wet_w_f",
                            base1 = "none")
# dummy data for watershed-ripeaian wetland (pct_wet_wr_f)
levels(DFobsz_alternate_dummy$pct_wet_wr_f)
df_dum_wet_wr = category_break(df_dum_wet$dat,
                               df_dum_wet$alt_dat,
                               "pct_wet_wr_f",
                               base1 = "none")
# dummy data for catchment wetland levels (pct_wet_c_f)
levels(DFobsz_alternate_dummy$pct_wet_c_f)
df_dum_wet_c = category_break(df_dum_wet_wr$dat,
                              df_dum_wet_wr$alt_dat,
                              "pct_wet_c_f",
                              base1 = "none")
# dummy data for catchment-riparian wetland levels (pct_wet_cr_f)
levels(DFobsz_alternate_dummy$pct_wet_cr_f)
df_dum_wet_cr = category_break(df_dum_wet_c$dat,
                               df_dum_wet_c$alt_dat,
                               "pct_wet_cr_f",
                               base1 = "none")

# dummy code for catchment-riparian impervious surfaces levels (pct_imp_rp_c_f)
levels(DFobsz_alternate_dummy$pct_imp_rp_c_f)
df_dum_imp_rp = category_break(df_dum_wet_cr$dat,
                               df_dum_wet_cr$alt_dat,
                               "pct_imp_rp_c_f",
                               base1 = "none")

# dummy code for total phosphorus levels (tp_f)
levels(DFobsz_alternate_dummy$tp_f)
df_dum_tp = category_break(df_dum_imp_rp$dat,
                           df_dum_imp_rp$alt_dat,
                           "tp_f",
                           base1 = "low")

dat_dummy = df_dum_tp$dat
alt_dat_dummy = df_dum_tp$alt_dat

```

## TLH Chunk 2 Bestglm and SSN0 Nugget Model-Travis

The code below evaluates if a different reference group was chosen would a qualitatively different model result from the best glm selection. For example, if the reference watershed used was vahusb_JM instead of vahusb_JU.

```{r ws_bestglm-Travis}

#double checking our base cases for variables we're selecting for the bestglm
df_dum_vahusb$base1
df_dum_vahusb$base2
df_dum_tp$base1
df_dum_tp$base2
# initial selection
WsWq <- dat_dummy|>
  dplyr::select(precip_mm, tmean, pct_imp_w_emplog, pct_for_w_emplog, pct_hay_w_emplog, pct_grs_w_emplog, elev_ws, do, p_h, tothab, l_spc, l_tds, l_turb, l_tn, vahusb_JU, vahusb_JR, vahusb_JA, vahusb_JL, bin_wet_w, tp_f_medium, tp_f_high)
str(WsWq)
# X can only contain numeric or factor
WsWq<- apply(WsWq, 2, as.numeric)

WsWqy<-cbind.data.frame(WsWq,y=DFobsz$vsci)
ASRcum_wswq1.best<-bestglm(WsWqy,IC = "AIC", nvmax =4, TopModels=15)

# See what terms are in best models
ASRcum_wswq1.best$BestModels
ASRcum_wswq1.best$Subsets
summary(ASRcum_wswq1.best$BestModel)

wswq_mlry2 <- lm(vsci ~ pct_imp_w_emplog + elev_ws + do + vahusb_JL, data = DFobsz)
summary(wswq_mlry2)
check_model(wswq_mlry2)
# if you wish to view avPlots uncomment below. But this hangs if you attempt to run the whole script as one block
#avPlots(wswq_mlry2)

## testing alternate base case to see if any issues arise
WsWq2 <- alt_dat_dummy|>
  dplyr::select(precip_mm, tmean, pct_imp_w_emplog, pct_for_w_emplog, pct_hay_w_emplog, pct_grs_w_emplog, elev_ws, do, p_h, tothab, l_spc, l_tds, l_turb, l_tn, vahusb_JM, vahusb_JR, vahusb_JA, vahusb_JL, bin_wet_w, tp_f_low, tp_f_high)

WsWq2<- apply(WsWq2, 2, as.numeric)

WsWqy2<-cbind.data.frame(WsWq2,y=DFobsz$vsci)
ASRcum_wswq2.best<-bestglm(WsWqy2,IC = "AIC", nvmax =4, TopModels=15)
summary(ASRcum_wswq1.best$BestModel)

# Perfect we get the same model we had before, no issues there.
```
